---
title: "Installing Rtools and Hector for Windows"
author: "Joe Brown"
date: "2023-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What is Rtools

Rtools is a toolchain that allows for users to build R packages from source code. Many R users may never need to interact with Rtools, however some tools are written in C++ or another programming language and thus need to be compiled into a format that can be executed in R. This is where Rtools comes in handy. With Rtools, R users can easily install and use R packages that require compilation from source.

Rtools is primarily used in the Windows environment. This vignette will briefly walk through the steps of installing Rtools on your Windows machine. We will conclude this vignette with a Hector installation to test that Rtools is working properly.

[See the video linked here for additional help!](https://www.youtube.com/watch?v=Pg3WY8wg1cY&t=457s)

# Downloading and installing Rtools

[HERE](https://cran.r-project.org/bin/windows/Rtools/) you can find links to downloads for different versions of Rtools. It is important that you download and install the RTools version that is compatible with whatever version of R you are working with.

## Installing Rtools

1.  Navigate to the folder where the Rtools installer was downloaded (mine is in downloads).

2.  Click the installer to run. Ensure that you note your destination location. It may be best to keep the default suggestion.

3.  When you see "Additional Tasks" window, make sure both boxes are checked. Click the next button to continue.

4.  When you see that "Ready to Install" window, click "Install" to begin the installation.

5.  Click "Finish" when installation is complete.

# Checking Installation

Before proceeding with building an R package from source, we want to check that the installation was successful. 

We can do this using `Sys.which()` in RStudio:

```{r}
Sys.which("make")
```
When the code above returns the path of your Rtools `make.exe`, you know that you have completed a successful installation and can proceed with compiling R packages from source.




